\section*{Chapter 5: Reducibility}
\addcontentsline{toc}{section}{Chapter 5: Reducibility}

\subsection*{5.2}

Show that EQ CFG is co-Turing-recognizable.


\subsection*{5.5}

Show that ATM is not mapping reducible to ETM . In other words, show that no computable function reduces ATM to ETM . (Hint: Use a proof by contradiction, and facts you already know about ATM and ETM .)


\subsection*{5.9}

Let AMBIGCFG = \{hGi| G is an ambiguous CFG\}. Show that AMBIGCFG is undecidable. (Hint: Use a reduction from PCP. Given an instance $\langle t_1, b_1\rangle, \langle t_2, b_2\rangle, \ldots, \langle t_k, b_k\rangle$ of the Post Correspondence Problem, construct a CFG G with the rules

S $\to$ T |B T $\to$ t1 T a1 | · · · | tk T ak | t1 a1 | · · · | tk ak B $\to$ b1 Ba1 | · · · | bk Bak | b1 a1 | · · · | bk ak ,

where a1 , . . . , ak are new terminal symbols. Prove that this reduction works.)


\subsection*{5.13}

Give an example of an undecidable language B, where B $\le$m B.


\subsection*{5.17}

Show that both conditions in Problem 5.16 are necessary for proving that P is undecidable.


\subsection*{5.21}

Consider the problem of determining whether a PDA accepts some string of the form \{ww| w $\in$ \{0,1\}$\ast$ \} . Use the computation history method to show that this problem is undecidable.


\subsection*{5.22}

Let X = \{hM, wi| M is a single-tape TM that never modifies the portion of the tape that contains the input w\}. Is X decidable? Prove your answer.


\subsection*{5.24}

Say that a CFG is minimal if none of its rules can be removed without changing the language generated. Let MIN CFG = \{hGi| G is a minimal CFG\}.

\begin{enumerate}[label=(\alph*), leftmargin=2.5em]

\item Show that MIN CFG is T-recognizable.

\item Show that MIN CFG is undecidable.

\end{enumerate}


\subsection*{5.28}

Consider the problem of determining whether a single-tape Turing machine ever writes a blank symbol over a nonblank symbol during the course of its computation on any input string. Formulate this problem as a language and show that it is undecidable.


\subsection*{5.33}

Show that the Post Correspondence Problem is decidable over the unary alphabet $\Sigma$ = \{1\}.


\clearpage
