\section{Chapter 9: Intractability}

\subsection*{9.2}

Prove that TIME(2n ) ( TIME(22n ).


\subsection*{9.7}

Give regular expressions with exponentiation that generate the following languages over the alphabet \{0,1\}.

\begin{enumerate}[label=(\alph*), leftmargin=2.5em]

\item All strings of length 500

\item All strings of length 500 or less

\item All strings of length 500 or more

\item All strings of length different than 500

\item All strings that contain exactly 500 1s

\item All strings that contain at least 500 1s

\item All strings that contain at most 500 1s

\item All strings of length 500 or more that contain a 0 in the 500th position

\item All strings that contain two 0s that have at least 500 symbols between them

\end{enumerate}


\subsection*{9.8}

If R is a regular expression, let R\{m,n\} represent the expression

Rm $\cup$ Rm+1 $\cup$ · · · $\cup$ Rn .

Show how to implement the R\{m,n\} operator, using the ordinary exponentiation operator, but without "· · · ".


\subsection*{9.9}

Show that if NP = PSAT, then NP = coNP.


\subsection*{9.12}

Recall that you may consider circuits that output strings over \{0,1\} by designating several output gates. Let add n : \{0,1\}2n -$\to$\{0,1\}n+1 take two n bit binary integers and produce the n + 1 bit sum. Show that you can compute the add n function with O(n) size circuits.


\subsection*{9.16}

Prove that an oracle C exists for which NPC $\neq$ coNPC .


\subsection*{9.19}

Let EREX↑ = \{hRi| R is a regular expression with exponentiation and L(R) = $\varnothing$\}. Show that EREX↑ $\in$ P.


\subsection*{9.23}

Define pad as in Problem 9.21.

\begin{enumerate}[label=(\alph*), leftmargin=2.5em]

\item Prove that for every A and natural number k, A $\in$ P iff pad (A, nk ) $\in$ P.

\item Prove that P $\neq$ SPACE(n).

\end{enumerate}


\subsection*{9.25}

Read the definition of a 2DFA (two-headed finite automaton) given in Problem 5.14. Prove that P contains a language that is not recognizable by a 2DFA.


\clearpage
